Discrete Quantum Theories and Computing

### 2018-04-06 dissertation.tex
dissertation.lyx "Chapter 1 Introduction" Updating...

Before diving into quantum computing, it is instructive to review
two kinds of classical computers: digital and analog. Although discrete
digital computers could be modeled by various computational models,
including Turing machines~\cite{Turing_1937}, these models are essentially
the same by the (strong) Church-Turing thesis \cite{Church_1936,Turing_1937,Kleene_1943}.
Moreover, any Turing-computable function could be carried out by a
digital computer as long as it has enough resources~\cite{Piccinini2015}.
In contrast, there are various idealized real computation models \cite{Siegelmann1998,Ziegler2007,weihrauch2012computable,blum2012complexity}.
Although these models are usually more powerful than discrete Turing
machines, their computational powers might not be realized by realistic
analog chips and computers, and the realized analog computational
devices are mostly chips embedded in our digital computers to communicate
with non-computational analog devices~\cite{Camenzind2005}. Currently
most of the quantum computing models based on the continuum of real
numbers. Similar to the development classical computers, the computational
power of realistic quantum computers might not be the same as predicted
by current quantum computing models. When developing realistic quantum
computers or estimating the computational powers of realistic quantum
computers, it might also be helpful to study discrete quantum computing
models.

Despite the history of computation, when replacing the continuum of
complex and real numbers by the discrete finite fields and finite
number of intervals, we show how a number of subtle properties of
quantum computing can be teased apart, step by step, as we explore
the implications of \emph{discrete quantum theories} in a systematic
fashion. In particular, on different discrete quantum theories, we
will present how geometrical structure of states and quantum probability
could be, and how these mathematical structures affect the power of
discrete quantum computing.

>>>>

This marriage of quantum mechanics and computer science first envisioned
and popularized by Feynman has created an awkward, but opportune,
moment. The embarrassing dilemma was concisely described by Aaronson
with the following three statements \cite{AaronsonArkhipov2011,BFRDARW2013}:
\begin{enumerate}[label=(\roman{enumi})]
\item \label{enu:Textbook-quantum-mechanics}Textbook quantum mechanics
is correct. 
\item \label{enu:no-efficient-classical-factoring}There does not exist
an efficient classical factoring algorithm. 
\item \label{enu:extended-Church-Turing-Thesis}The extended Church-Turing
thesis — that probabilistic Turing machines can efficiently simulate
any physically realizable model of computation — is correct \cite{BernsteinVazirani1997,Kaye2007}. 
\end{enumerate}
There is overwhelming evidence to support each of these statements.
The theoretical framework of quantum mechanics \ref{enu:Textbook-quantum-mechanics}
has withstood decades of experimental confirmation. Entire industries
are founded on the assumption \ref{enu:no-efficient-classical-factoring}
that algorithms like RSA are secure and they also have withstood years
of attempted attacks \cite{boneh1999twenty,wiki:RSAFactoring}. Finally
the entire field of complexity theory in computer science which has
also withstood years of field testing rests, in essence, on assumption
\ref{enu:extended-Church-Turing-Thesis} \cite{Aaronson2005,Piccinini2015}.
And yet at least one of these three statements \emph{must be false}!
Indeed if there is a corresponding efficient classical factoring algorithm,
then we concede \ref{enu:no-efficient-classical-factoring}. If it
is correct Shor's efficient factoring algorithm is realizable, and
we can prove there is no efficient classical factoring algorithm,
we concede \ref{enu:extended-Church-Turing-Thesis}. Otherwise, if
we cannot implement Shor's algorithm no matter how hard we try, textbook
quantum mechanics, i.e., \ref{enu:Textbook-quantum-mechanics} may
need to be improved by a better theory. It is unlikely that there
will be a simple resolution to this awkward situation. It is more
likely that the resolution will emerge from deep and careful analyses
of the foundations of each field.

analog computing and probabilistic computing

In common practice, randomized algorithms are approximated using a
pseudorandom number generator in place of a true source of random
bits; such an implementation may deviate from the expected theoretical
behavior~\cite{wiki:RandomizedAlgorithm}.
